evidence of adaptation-independent time-warped spike timing noise correlations in cortical networks

### 2020-08-11 manuscript.tex
gut

Such correlated clusters were observed in a large number of first spike pair distributions (Supplementary Vid. 1) and represent reliable regions of the response distribution space that can be analysed to test the nature of spike timing noise correlations. The correlative structure could be caused by neural adaptation, a continuous non-adaptative latent state or a combination of the two.

@@

A novel algorithm extracted isolated clusters from single stimulus first spike pair response distributions, demonstrating that temporally precise spike patterns fire on subsets of trials, as shown previously for the data set \cite{reyes2014laminar}. The algorithm, however, also showed that first spike pairs within these isolated clusters were positively correlated above chance-levels. A large proportion of these noise correlations would have been missed without clustering. Moreover, many of the correlated clusters had non-unitary correlation angles.

Importantly, clusters remained correlated with non-unitary gradient after controlling for neural adaptation,
% using time series analysis, 
giving evidence of precise non-adaptive noise correlations. Such non-unitary spike time correlation angles cannot be explained by a trial-to-trial shift in the onset of the stimulus or cortical activity, which would produce correlations with a unitary gradient (i.e. a $45^{\circ}$ angle). Instead the `non-unitary' correlation gradients may be indicative of covariation by a non-adaptative continuous latent state that varies from trial-to-trial.

This novel finding of `adaptation-independent clustered time-warped spike-timing noise correlations' shows that the distribution of decodable first spike time relative differences for pairs of neurons are `warped' from trial-to-trial by some `non-adaptive' continuous latent state.
% This phenomenon is referred to as `adaptation-independent clustered time-warped spike-timing noise correlations'. 
This differs from previous `time-warp' studies, which aim to align spike trains over trials and do not simultaneously consider stimulus encoding, clustering, adaptation and different warping functions for different neurons \cite{williams2020discovering}.

>>>>

Correlated clusters were observed in a large number of first spike pair distributions (Supplementary Vid. 1). A novel algorithm extracted isolated clusters from single stimulus first spike pair response distributions, demonstrating that temporally precise spike patterns fire on subsets of trials, as shown previously for the data set \cite{reyes2014laminar}. The algorithm, however, also showed that first spike pairs within these isolated clusters were positively correlated above chance-levels. These correlated clusters represent reliable regions of response distribution space that allow analysis of spike timing noise correlations possibly caused by neural adaptation, a continuous non-adaptative latent state or a combination of both. A large proportion of these noise correlations would have been missed without clustering.

@@

Interestingly, many of the correlated clusters had non-unitary correlation angles. 
Importantly, clusters remained correlated with non-unitary gradient after controlling for neural adaptation,
% using time series analysis, 
giving evidence of precise non-adaptive noise correlations. Such non-unitary spike time correlation angles cannot be explained by a trial-to-trial shift in the onset of the stimulus or cortical activity, which would produce correlations with a unitary gradient (i.e. a $45^{\circ}$ angle). Instead the `non-unitary' correlation gradients are indicative of covariation by a non-adaptative continuous latent state that varies from trial-to-trial.

This novel finding of `adaptation-independent clustered time-warp spike-timing noise correlations' shows that the distribution of decodable first spike time relative differences for pairs of neurons are `warped' from trial-to-trial by some `non-adaptive' continuous latent state.
% This phenomenon is referred to as `adaptation-independent clustered time-warped spike-timing noise correlations'. 
This differs from previous `time-warp' studies, which aim to align spike trains over trials and do not simultaneously consider stimulus encoding, clustering, adaptation and different warping functions for different neurons \cite{williams2020discovering}.
